---
title: "Simulation results"
output: 
  html_document:
    number_sections: FALSE
    toc: true
    toc_float: true
    theme: united
date: "2025-08-28"
---

```{r setup, include=FALSE}
# load in packages 
library(gamlss)
library(gamlss.dist)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## The setup 

I'm using this just to show the results, together with the diagnostics, graphs and all the good stuff.

<br>

*Table 1: Simulations proposal of different changing parameters*

|       | Mean         | Variance     | Skewness     | Kurtosis     |
|-------|--------------|--------------|--------------|--------------|
| Sim 1 | constant     | time-varying | \-           | \-           |
| Sim 2 | time-varying | time-varying | \-           | \-           |
| Sim 3 | time-varying | time-varying | constant     | \-           |
| Sim 4 | time-varying | time-varying | \-           | constant     |
| Sim 5 | time-varying | time-varying | constant     | constant     |
| Sim 6 | time-varying | time-varying | time-varying | constant     |
| Sim 7 | time-varying | time-varying | constant     | time-varying |
| Sim 8 | time-varying | time-varying | time-varying | time-varying |

<br> 

Other things to consider: 

1. Changing temporal resolution: match the resolution of the Helgoland data, then we could down-sample to see how resolution affects recovery 
2. 


<br>

## Simulation 1

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
